---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
```


# GuanRankR

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->


## Overview

An R Package for computing [Guan Rank](https://web.cse.ohio-state.edu/~zhang.10631/files/survival_2021.pdf), 
which transforms survival data in to an easy to model ranking

The method was used in many of [Prof. Yuanfang Guan's](https://medicine.umich.edu/dept/dcmb/yuanfang-guan-phd) top performing 
[DREAM Challenge models](https://www.synapse.org/#!Synapse:syn6180942/wiki/401735)

## Installation


Install the released version of GuanRankR from Github:

```{r}
remotes::install_github(repo = "Systems-Methods/GuanRankR")
```

Or:

```{r}
remotes::install_git(
  'https://github.com/Systems-Methods/GuanRankR.git'
)
```


Or install the development version

```{r}
remotes::install_github(repo = "Systems-Methods/GuanRankR", 
                        ref = "develop")

```


## Example

### Example Data

First we can use the [{UCSCXenaTools}](https://cran.r-project.org/web/packages/UCSCXenaTools/vignettes/USCSXenaTools.html) 
package to download an example mRNASeq dataset.

```{r, eval = TRUE}
dat <- UCSCXenaTools::getTCGAdata(project = "LUAD", 
                                  clinical = TRUE, 
                                  download = TRUE,
                                  quiet = TRUE)

clin <- data.table::fread(dat$destfiles, data.table = FALSE)
```

### Create Survival Data and Run Ranks

```{r, eval = TRUE}

survData <- data.frame(
  time = ifelse(!is.na(clin$days_to_last_followup),
                as.numeric(clin$days_to_last_followup),
                as.numeric(clin$days_to_death)),
  status = dplyr::recode(
    clin$vital_status, 'LIVING' = 0,'DECEASED' = 1, .default = NA_real_
  )
)

gr <- GuanRankR::calculate_guan_rank(surv_data = survData)


```

### Plotting Results


```{r, eval = TRUE}
Col <- as.vector(factor(gr$status, labels = c('slategrey', 'cyan')))
par(mfrow = c(2,2))
plot(survival::survfit(survival::Surv(gr$time, gr$status)~1), xlab = "Days",
     ylab = "Survival %", conf.int = FALSE, mark.time = TRUE, col = "slategrey",
     cex = .5, main = 'LUAD')
plot(gr$time,gr$guan_rank, pch = 19, col = Col, cex = .5, xlab = "Days",
     ylab = "Guan Rank", main = "Guan Rank vs Time")
boxplot(gr$guan_rank ~ gr$status, xlab = "Status", ylab = "Guan Rank",
        main = "Guan Rank vs Status")
hist(gr$guan_rank, breaks = 30, xlab = "Guan Rank",
     main = "Guan Rank Distribution")

```


## Code of Conduct
  
Please note that the GuanRankR project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
